---
title: "Ratio Analysis"
output: pdf_document
date: "2025-12-03"
---

```{r, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ratio Table

The data table summarizes several measles outbreaks from different states and years.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(readr)
library(knitr)
library(metafor)
df <- read_csv("RawData/measles_summary.csv")

kable(df, caption = "Measles Outbreak Summary Table")


```

## Meta-Analysis of Ratios

I used a meta analysis to combine the ratios from all outbreaks into one overall estimate. Meta analysis is a statistical method that combines results from multiple datasets or studies into one pooled estimate.

Meta-analysis gives one combined ratio that:

* gives more weight to larger outbreaks

* gives less weight to smaller outbreaks

* gives a confidence interval (range of plausible values)

```{r, echo=FALSE}
### ---------------------------
### Meta-analysis of ratios
### ---------------------------

# Adult (>=18) cases for each outbreak
A <- df$Cases_above_18  

# Child (5–17) cases
C <- df$Cases_5_17

# Names of outbreaks
outbreaks <- df$`Location/State`

# Compute ratios
R <- A / C

# Log ratios
y <- log(R)

# Standard error of log-ratio: sqrt( 1/A + 1/C )
SE <- sqrt(1/A + 1/C)

# Variances
v <- SE^2


res_re <- rma.uni(
  yi = y,
  vi = v,
  method = "REML"   # Random-effects model using REML estimator
)
# ------------------------------------------------------------
# 4. Print detailed model results
# ------------------------------------------------------------
print(res_re)

# ------------------------------------------------------------
# 5. Convert pooled estimate back to ratio scale
# ------------------------------------------------------------
pooled_ratio_re <- exp(res_re$b)
ci_lower_re     <- exp(res_re$ci.lb)
ci_upper_re     <- exp(res_re$ci.ub)

cat("\n----------------------------------------\n")
cat("Random-Effects Pooled Ratio:", round(pooled_ratio_re, 3), "\n")
cat("95% CI:", round(ci_lower_re, 3), "–", round(ci_upper_re, 3), "\n")
cat("Tau^2 (between-outbreak variance):", round(res_re$tau2, 4), "\n")
cat("----------------------------------------\n\n")

```


## Final result

In a random effects meta analysis of six outbreaks, the pooled 18+ to 5-17 years case ratio was **0.995 (95% CI: 0.73–1.37)**, indicating no consistent overall tendency for either adults or 5–17 year olds to have higher measles case rates on average across outbreaks. However, high heterogeneity (I² = 90%, $\tau^2$ = .13), suggesting that the ratio varies widely between outbreaks.




